P1, C1.1, delete sentence This thesis…animals

DONE

P1, C1.1 delete para This thesis..function


P2 C1.1 delete “that will be discussed…chapters”


DONE

P2 C1.1.1 (see figure 1.2) - just (figure 1.2) , just to be consistent with other fig references.


DONE
P3 C1.1.1 para The term…[149] - suggest move this up a bit, its a historical introduction, maybe make it second paragraph in this section?

DONE

P3 C1.1.2 para 1. Quite recently ->until recently

DONE

P4 “high level functional the synapse” - again these two paragraphs “the synapse…between neurons” feel like good introduction and maybe needs to come before some of the text above. Again maybe second paragraph in this section?

DONE and removed High level function of synapse subsubsection

P4 Figure 1.2 “Picture credit” suggest chage this to either “Taken from:” or “Modified from:”

DONE

P4 C1.2 “gap” - just call it a cleft since thats what you use thereafter?

DONE

P5 1.2.1. Have a read through this again. I’m not sure the order of the sentences really works and might be better in a slightly different order to make more sense.


Edited order should be
Presynapse and gross features and development
Contain vesicles these are docked to active zone
Molecular biology of active zone
Information processing and post activation changes

P6 “They are relatively” - what is actually resistant? the protein or the network? Reads like the protein in current context….

It is TODO 
P6 1.2.3 Calcium ion…depression” delete

DONE
P7 4 sub-units - four sub-units

DONE
P7 notes to remove

DONE
P10 1.2.6 - The first sentence does not make sense and actually could just be deleted. then The PSP has ancient origins: The proteins evolved around the time of the earliest…

DONE

P11 1.2.9- I think this is written backwards. Don’t you need to introduce what a PPI is first before going into database quality? Also is it really the database quality or the interaction data quality?. The first paragraph os 1.2.10 would be a good introduction.

changed to

Databases include high quality experimental evidence of protein interactions and use standardised ontologies to organise their information \cite{brazma2006standards}, \cite{hermjakob2004hupo}, \cite{kerrien2007broadening}. Databases contain evidence on interactions that varies in reliability from direct experimental records to inferred interactions (e.g. literature data mining and co-expression).


P13 "To explain why…disorders” delete

DONE

P14 question. I always put quotes in both italics and quotation marks so it is ultra clear.

DONE footnote removed 
P16 1.3.4& general from now on. More X-refs. You mentions SNPs here for the first time but you have a good description later  in 1.4.1, stick in a cross reference here.
DONE
todo removed
P17 1.3.4 MAGMA (as above, cross ref to your section on MAGMA
DONE CROSS REF ADDED TO MAGMA GSA

P19 1.3.6 “I will…educational attainment” - suggest this is edited down and moved to the end of chapter 1 as part of the hypothesis and how you are going to test it?

MOVED

P20-P28 - this section feels a bit methods but probably best where it is since chapter 2 has a different set of aims (aims in terms of your thesis)
P28 figure 1.3 - is this figure useful or interesting?

NO REMOVED

P30 Plan
I would expand this a little. Your reader has just spent pages on GWAS, GSA, MAGMA etc. Bring them back to the idea that what you propose is that we can use synapse networks built from proteomics and PPIs and use clustingg to derive new communities which might be useful gene sets. Then just clearly state your hypothesis. Then briefly explain how you plan to test your hypothesis and maybe even explain what a positive and or negative result would look like.. Once you have done that you don’t need to introduce the following chapters, they just make sense.

P31 general - you have “results” subheadings in multiple places - are these really results or are they more accurately  "validation of datasets and/or methods used"
P31 - intro chapter doesn’t really add value as it is. Need to explain that you are going to explain where the datasets come from and included in that, in particular for the GWAs, demonstrate that the methodology you want to apply can replicate the previous findings - validating your methods and the data.. Essentially this is all due diligence since it is largely stuff from other people.
P31 2.1 Chapter title(and others) are a bit terse. Network graph of what?

DONE of the post synaptic prtoeme
P34 figure??

DONE
P35 and elsewhere in this chapter. “The fine;l filtered list contained 325….” - where does the reader get a copy of this list?

This was from Katharinas PhD
P36 “the original gal file” - again is this provided somewhere as supplementary material or on GitHub? supplementary material

Supplemental
On the whole chapter 2 is good, is it worth a mini conclusion at the end of the summary that on the basis of the checks performed you have the confidence to proceed in the next chapters which focus on your development and testing of the hypotheses from chapter 1?

DONE

I've removed some of the subscripts in titles eg Education\textsubscript{Discovery} as it is handled as a warning by latex 